### 2012-03-07 abstract.tex
Begin work on abstract.

Nell is targeted at children far from educational infrastructure.

We present Project Nell, tablet software to teach reading and writing to
children far from educational infrastructure.  Nell is a modular
narrative system with ``a low floor and no ceilings'': the design
scales from teaching letter shapes to programming the system itself.

Nell is the successor to the Sugar educational software developed by
One Laptop per Child for their XO-1 laptop.  Nell targets the XO-3
tablet, a sub-\$100 solar-powered tablet for the developing world.
Experience with Sugar drove design choices for Nell.

\footnote Nell was inspired by the ``Young Ladies' Interactive
Primer'' in the Neal Stephenson novel, ``The Diamond Age''.  It takes
its name from novel's protagonist, but could also be read as an
acronym for ``Narrative Environment for Learning Learning'' (a nod to
Seymour Papert).